---
output:
  pdf_document:
    toc: false
    number_sections: false
---

# Graphical Analysis: Productivity Differences and Export Flows to China (2005)

# 1. R code used for the analysis

The following code loads the gravity data, restricts the sample to year 2005 with China as the only importer (destination), and constructs a measure of productivity difference (absolute difference in log GDP per capita between exporter and China). It then plots log export flow to China against this productivity difference and fits a linear regression.

```{r setup-and-code, echo = TRUE, message = FALSE, warning = FALSE}
# Path to data (edit if your gravity.csv is elsewhere)
FOLDER <- "c:/Users/dylan/Downloads/gravity"
csv_path <- file.path(FOLDER, "gravity.csv")
df <- read.csv(csv_path, stringsAsFactors = FALSE)

# Restrict: year 2005, China as importer only
df <- df[df$year == 2005 & df$iso_d == "CHN", ]
df <- df[complete.cases(df[, c("flow", "gdpcap_o", "gdpcap_d")]), ]
df <- df[df$flow > 0 & df$gdpcap_o > 0 & df$gdpcap_d > 0, ]

# Productivity difference: |log(gdpcap_exporter) - log(gdpcap_China)|
df$log_gdpcap_diff <- abs(log(df$gdpcap_o) - log(df$gdpcap_d))
df$log_flow <- log(df$flow)

# Linear regression
fit <- lm(log_flow ~ log_gdpcap_diff, data = df)
slope <- coef(fit)["log_gdpcap_diff"]
intercept <- coef(fit)["(Intercept)"]
r_sq <- summary(fit)$r.squared
n_obs <- nrow(df)
```

# 2. Graph

Scatter plot of log export flow to China (vertical axis) against the absolute log GDP per capita difference between exporter and China (horizontal axis), with fitted regression line.

```{r graph, echo = FALSE, fig.width = 6.5, fig.height = 5}
par(mar = c(4, 4, 3, 1))
plot(
  df$log_gdpcap_diff,
  df$log_flow,
  pch = 16,
  col = adjustcolor("darkblue", alpha.f = 0.6),
  cex = 1.2,
  xlab = "Absolute log GDP per capita difference (exporter vs China)",
  ylab = "Log export flow to China",
  main = "Export flows to China (2005): productivity difference and bilateral trade"
)
abline(fit, col = "red", lwd = 2)
legend(
  "topright",
  legend = sprintf("Linear fit: slope = %.3f, R² = %.3f", slope, r_sq),
  col = "red",
  lty = 1,
  lwd = 2
)
grid()
```

# 3. Written explanation of the results

**Sample.** The analysis uses the gravity dataset restricted to **year 2005** with **China as the only importer** (destination). Thus each observation is a country that exports to China. The sample includes only exporter–China pairs with strictly positive export flow and non-missing GDP per capita for both the exporter and China, yielding `r n_obs` observations.

**Variables.** The vertical axis is **log export flow to China** (natural log of bilateral export value from the exporter to China). The horizontal axis is the **absolute difference in log GDP per capita** between the exporter and China, |ln(GDP per capita_exporter) − ln(GDP per capita_China)|. This is used as a proxy for **productivity (or income) difference**: larger values mean the exporter and China are more different in terms of income per capita.

**Finding.** The scatter plot shows the relationship between this productivity difference and log export flow to China. The red line is the ordinary least squares (OLS) regression of log export flow on the productivity-difference measure.

* **Slope:** The estimated slope is **`r round(slope, 3)`**. If the slope is **positive**, then, on average, larger productivity differences between the exporter and China are associated with **higher** export flows to China. That is consistent with comparative-advantage or gravity-style reasoning: countries that are more different in productivity may trade more with China, all else equal. If the slope is **negative**, then larger productivity differences are associated with **lower** export flows to China in this cross-section.

* **Fit:** The regression has an R² of **`r round(r_sq, 3)`**, so the linear relationship explains about `r round(100 * r_sq, 1)`% of the variation in log export flow to China. The remainder is explained by other factors (e.g. distance, size, trade policy, and unobserved heterogeneity).

**Conclusion.** This graphical analysis indicates whether, in 2005, **differences in productivity (proxied by the absolute log GDP per capita difference) are associated with higher export flows to China**. The sign and magnitude of the slope summarize that relationship; the graph allows one to check for nonlinearity or outliers. For a fuller causal interpretation, one would need to control for confounding variables (e.g. distance, contiguity, RTAs) in a multivariate gravity regression.
